---
title: "Check the 2020 data and incorporate it into gfiphc"
author: "Andrew Edwards"
output: pdf_document
fontsize: 12pt
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ">",
  fig.width = 10,
  fig.height = 8
)
```

In 2020 only the first 20 hooks were evaluated, so those data are not easily imported into
GFbio. Going to incorporate into gfiphc here. Likely need this as a template for
future years (resave this file with new year, and change all 2020's to the
subsequent year -- and go through the code somewhat manually to check the output
as you go along). This code includes some
manual checks to make sure the data look
okay.

For comparison first look at 2013 data included in gfiphc:
<!-- Based on iphc2013data.Rnw -->
```{r setup}
load_all()
setData2013
countData2013
```

We want to get the new data into the same format as those (columns with same
names and classes, even though in retrospect some classes aren't ideally
chosen, but also retaining retrieved and observed hooks for the set data). Two
data sets are
needed because later gfiphc code summarises catches of a particular species at the
station level, and needs to create counts of zeros for the species of interest
(and such zeros are not included in IPHC output).

## Set-level information

Maria was sent the file `2020 IPHCtoDFO_dataExtraction-Maria.xls` for set
details, but this is multiple sheets and more complex than needed. Will try
extracting directly from IPHC website (which they want us to do in the future
anyway), using the following instructions:

Go to https://www.iphc.int/data/fiss-data-query and select the following options:

1. Year Range -- 2020 to 2020.

2. Area 2B

3. Purpose Codes -- All

4. IPHC Charter Regions -- All

5. Maps -- Nothing

6. Select non-Pacific halibut species -- deselect All.

Download tab on bottom right (see instructions above question 4), and select
CrossTab. Select "Set and Pacific Halibut data" and .xlsx format (I tried .csv
format but it didn't save with commas, strangely). Save in this
folder as `set-and-halibut-data-2020.xlsx`. Open in Excel and Export as .csv, `set-and-halibut-data-2020.csv`,
and when trying to quit Excel say no to save changes (not sure if that matters).

Repeat but with all non-halibut data (select All in number 6), and save as
`non-halibut-data-2020.xlsx` and export as .csv in Excel,
`non-halibut-data-2020.csv`. Importantly, this file
(but not the first one) contains the numbers of observed hooks, needed in our
calculations.

```{r loadsetdata}
sets_raw <- readr::read_csv("set-and-halibut-data-2020.csv") %>%
  dplyr::mutate_if(is.character, factor)
sets_raw
summary(sets_raw)
testthat::expect_equal(unique(sets_raw$"IPHC Reg Area"), as.factor("2B")) # Check just BC
testthat::expect_equal(unique(sets_raw$Year), 2020)
#testthat::expect_equal(length(unique(sets_raw$Station)),
#                       length(sets_raw$Station))
```

### Understand any issues raised above

Uncomment those three `testthat` commands when looking at new data each year. If
any of fail then have to comment it out and figure out what it means here.

This is for 2020 (check for future years), to look for station(s) that was
fished twice:
```{r understand}
length(unique(sets_raw$Station))
length(sets_raw$Station)
dplyr::count(sets_raw, Station) %>% dplyr::filter(n > 1)
twice_fished <- dplyr::count(sets_raw, Station) %>%
  dplyr::filter(n > 1) %>%
  select(Station) %>%
  as.numeric()
# If there's more than a single station then adapt later code
as.data.frame(dplyr::filter(sets_raw,
                            Station == twice_fished))

```
So Station `r twice_fished` had two vessels fishing the same station (which the code below
originally caused a total of four rows for that station, explaining the 200 rows
I had in original `setData2020` before fixing the issue). Interestingly the
halibut catches were almost double for one vessel than the other (but were 6
days apart):

Note that one of those entries has 'Vessel code' HAN, but HAN only appears once
in the whole data set (as seen in `summary(sets_raw)` above:
```{r duplicate}
summary(sets_raw$"Vessel code")
```

So given we want to exclude one of the duplicates, makes sense to exclude
HAN. (Also, Dana mentioned some gear comparison studies for 2020).
Do this and simplify down to what's needed and rename, based on `iphc2013data.Rnw` (need
to include the 'purpose' column, unlike 2013):
```{r simplify}
sets_simp <- dplyr::filter(sets_raw, `Vessel code` != "HAN") %>%
  dplyr::select(year = Year,
                station = Station,
                lat = "MidLat fished",
                lon = "MidLon fished",
                avgDepth = "AvgDepth (fm)",
                skatesHauled = "No. skates hauled",
                effSkateIPHC = "Effective skates hauled",
                soakTimeMinutes = "Soak time (min.)",  # Joe might want
                usable = Eff,
                purpose = Purpose) %>%
  arrange(station) %>%
  dplyr::mutate(year = as.integer(year),
                station = as.character(station),
                avgDepth = as.integer(avgDepth),
                usable = as.character(usable))
sets_simp
```

For future years check the HAN issue and remove that first line if necessary
(especially if HAN is used in the survey).

Then change purpose to `standard` (`Y/N`) to match 2018 data (`Y` for the
standard grid). Here `purpose` takes three values, and we need to convert to
`standard`:
```{r purpose}
summary(sets_simp$purpose)

sets_simp_std <- dplyr::mutate(sets_simp,
                               standard_tmp = (purpose == "Standard grid"))

standard <- as.character(sets_simp_std$standard_tmp)  # to get the right length
standard[sets_simp_std$standard_tmp] = "Y"
standard[!sets_simp_std$standard_tmp] = "N"
length(standard)

sets_simp_std <- cbind(sets_simp_std,
                   standard) %>%
  as_tibble() %>%
  dplyr::select(-c("standard_tmp"))
summary(sets_simp_std)
```

## Look at data and show map to understand changing definition of standard station from 2018 to 2020.

It seems that the definition of 'standard grid' has changed from 2018 (when
first needed due to the expanded grid) to 2020. Simply equating them as above is
not sufficient.

This section figures out the problem and corrects it. However, to replicate the original
analysis we need to revert the correction. So doing that here and using that for
the original analysis, so make station 2343 standard again:
```{r revert}
hooks_with_bait_revert <- hooks_with_bait

# This should be commented out for 2021 survey analysis in iphc-2021-data.Rmd,
#  since the problem is fixed:
hooks_with_bait_revert$set_counts[hooks_with_bait_revert$set_counts$year == 2018 &
                           hooks_with_bait_revert$set_counts$station == 2343,
                           ]$standard = "Y"

filter(hooks_with_bait$set_counts, year == 2018, station == 2343) %>%
  as.data.frame()      # saved version
filter(hooks_with_bait_revert$set_counts, year == 2018, station == 2343) %>%
  as.data.frame()      # reverted version

```

TODO - that first filter line currently says Y for standard. Should be N when other
 code is re-run and `hooks_with_bait` .rda file (and others) updated.

Plotting four years, with crosses showing 'non-standard'. (2020 is coloured
different since no hooks with bait data yet, but the important bit is the crosses).
```{r maps}
sets_2020 <- sets_simp_std  # not the same structure as sets_2018, below

plot_iphc_map_panel(hooks_with_bait_revert$set_counts,
               sp = "Hooks with bait",
               years = 2017:2019,
               indicate_standard = TRUE)

plot_iphc_map(sets_2020,
              sp = NULL,
              years = 2020,
              indicate_standard = TRUE)
```

Clearly 2020 has a few less stations just north of Vancouver Island, but not
enough to worry about greatly. The curious ones are the ones way in in the
inlets that are no longer flagged as non-standard.

Need to look and plot values:
```{r standardornot}
sets_2018 <- filter(hooks_with_bait_revert$set_counts,
                    year == 2018)
not_std_2018 <- filter(sets_2018,
                       standard == "N")$station


not_std_2020 <- filter(sets_2020,
                       standard == "N")$station

# Not standard in both:
not_std_2018_and_2020 <- intersect(not_std_2018, not_std_2020)
not_std_2018_and_2020

length(not_std_2018)
length(not_std_2020)
length(not_std_2018_and_2020)

# 2018 has some east of the map, all non-standard:
filter(hooks_with_bait_revert$set_counts, year == 2018, lon > -124)$standard
nrow(filter(hooks_with_bait_revert$set_counts, year == 2018, lon > -124))

std_in_2018_but_not_std_in_2020 <- intersect(filter(sets_2018,
                                                    standard == "Y")$station,
                                             not_std_2020)
std_in_2018_but_not_std_in_2020

not_std_in_2018_but_std_in_2020 <- intersect(not_std_2018,
                                             filter(sets_2020,
                                                    standard == "Y")$station)
not_std_in_2018_but_std_in_2020

# setdiff(x, y) - elements in x but not in y
# setdiff(not_std_2018, not_std_2020) - but 2020 fewer coverage so misleading
```

Plot stations not standard in 2018 but standard in 2020, and vice versa, using
each years' lats and lons (to verify that they all still agree -- i.e., that station
numbers have consistent lats and lons), and show 2019 data to check no 'usual'
stations are non-standard in 2018 or 2020:

```{r mapstandard}
plot_BC()
points(lat~lon,
       data = filter(sets_2018,
                     station %in% not_std_in_2018_but_std_in_2020),
       col="red",
       pch = 19)

# Do the same but using 2020 station co-ordinates - should overlap:
points(lat~lon,
       data = filter(sets_2020,
                     station %in% not_std_in_2018_but_std_in_2020),
       col="blue",
       pch = 3)

# And show the single station std in 2018 but not 2020:
points(lat~lon,
       data = filter(sets_2018,
                     station %in% std_in_2018_but_not_std_in_2020),
       col="red",
       pch = 17)
points(lat~lon,
       data = filter(sets_2020,
                     station %in% std_in_2018_but_not_std_in_2020),
       col="blue",
       pch = 1,
       cex = 2)

# Now show all 2019 stations:
points(lat~lon,
       data = filter(hooks_with_bait_revert$set_counts,
                     year == 2019),
       col="darkgreen",
       pch = 0)

legend("bottomleft",
       legend = c("Not std 2018 but std 2020 (2018 co-ords)",
                  "Not std 2018 but std 2020 (2020 co-ords)",
                  "Std 2018 but not std 2020 (2018 co-ords)",
                  "Std 2018 but not std 2020 (2020 co-ords)",
                  "All 2019 stations"),
             pch = c(19, 3, 17, 1, 0),
             pt.cex = c(1, 1, 1, 2, 1),
             col = c("red", "blue", "red", "blue", "darkgreen"))
```

Check if the one standard station in 2018 but not in 2020 (not fished at all in 2019)
appears in any earlier years:
```{r checkearly}
filter(hooks_with_bait_revert$set_counts,
       station == std_in_2018_but_not_std_in_2020) %>%
  as.data.frame()
```
Was only fished in 2018 and 2020, so we should define it as non-standard.

So, the conclusions from this section are that we should:

1. Retain the 2018 definitions of standard stations, not the new ones defined in 2020:
```{r corrected}
sets_simp_std_corrected <- sets_simp_std

sets_simp_std_corrected$standard[sets_simp_std_corrected$station %in%
                                 not_std_in_2018_but_std_in_2020] <- "N"
# cbind(sets_simp_std$standard, sets_simp_std_corrected$standard) # to check
```

2. Define station 2343 as non-standard (over-riding original 2018 designation).

So check which functions that need changing, since they create a 'standard'
column. These do not need changing: `get_iphc_hooks()` and `get_iphc_skates_info`.

Then `get_iphc_sets_info()` does retuun `standard`, but the `standard`
designation is not saved in GFbio it is saved in `setDataExpansion` in `gfiphc`.
So just need to add a line in `IPHC-stations-expanded.R` and then re-save all
.rda files. Fixed that, now recreating all .rda files, as per the README.


## Species counts

First, get the species counts into the desired format (to match `countData2013` shown earlier):
```{r counts}
counts_raw <- readr::read_csv("non-halibut-data-2020.csv") %>%
  dplyr::mutate_if(is.character, factor)

counts_raw
summary(counts_raw)
testthat::expect_equal(unique(counts_raw$Year), 2020)
testthat::expect_equal(unique(counts_raw$SampleType), as.factor("20Hook"))

# This mismatches for 2020:
# testthat::expect_equal(length(unique(counts_raw$Station)),
#                        length(sets_raw$Station))

unique(counts_raw$"Species Name")
```

Note that halibut are not included in these counts, which I presume explains why
total number of counts for a station does not add up to `HooksObserved`.

Need to remove the HAN records for the twice-fished station, which turns out to
be set number 4 for station 2104:
```{r twicefished}
dplyr::filter(counts_raw, Station == twice_fished) %>%
  dplyr::select(c("Station", "Setno", "Species Name",
                  "Number Observed")) %>%
    as.data.frame()

dplyr::filter(sets_raw, Station == twice_fished)
```

So use that here to remove the species counts for that vessel (note that vessel
code is not in `counts_raw`)
```{r countcalcs}
dplyr::filter(counts_raw,
              Station == twice_fished & Setno == 4)

# So just keep these:
dplyr::filter(counts_raw,
              !(Station == twice_fished & Setno == 4))

countData2020 <- dplyr::filter(counts_raw,
                               !(Station == twice_fished & Setno == 4)) %>%
  dplyr::select(year = Year,
                station = Station,
                spNameIPHC = "Species Name",
                specCount = "Number Observed") %>%
  arrange(station) %>%
  dplyr::mutate(year = as.integer(year),
                station = as.character(station),
                spNameIPHC = as.character(spNameIPHC),
                specCount = as.integer(specCount))

testthat::expect_equal(names(countData2013), names(countData2020))
countData2020
summary(countData2020)

usethis::use_data(countData2020,
                  overwrite = TRUE)

```

## Hooks observed and retrieved

Now, obtain the numbers of hooks observed and retrieved from `counts_raw`, to
then merge into the set details:
```{r hookdetails}
hook_details <- dplyr::filter(counts_raw,
                              !(Station == twice_fished & Setno == 4)) %>%
  dplyr::group_by(Station) %>%
  dplyr::summarise(year = unique(Year),
                   hooksRetr = unique(HooksRetrieved),
                   hooksObs = unique(HooksObserved)) %>%
  dplyr::rename(station = Station) %>%
  dplyr::ungroup() %>%
  arrange(station) %>%
  dplyr::mutate(year = as.integer(year),
                station = as.character(station))

hook_details

expect_equal(sets_simp_std_corrected$station, hook_details$station)
```

So now need to get the hook details into the set details, and keep columns as
for `setData2013` but also with `standard`, and may as well keep `hooksRetr` and
`hooksObs`:
```{r merge}
setData2020 <- dplyr::left_join(sets_simp_std_corrected,
                                hook_details,
                                by = c("year", "station")) %>%
  dplyr::mutate(E_it20 = effSkateIPHC * hooksObs / hooksRetr) %>%
  dplyr::select(year,
                station,
                lat,
                lon,
                avgDepth,
                effSkateIPHC,
                E_it20,
                usable,
                standard,
                hooksRetr,
                hooksObs) %>%
  dplyr::mutate(year = as.integer(year),
                station = as.character(station),
                avgDepth = as.integer(avgDepth),
                usable = as.character(usable),
                standard = as.factor(standard))
setData2020
testthat::expect_equal(names(setData2013), names(setData2020)[1:ncol(setData2013)])
summary(setData2020)
usethis::use_data(setData2020,
                  overwrite = TRUE)
```

## Check species names

The file `inst/extdata/iphc-spp-names.csv` contains species common names (as
used for gfsynopsis, and a few extra like `unidentified skate`) and the IPHC
common name.

These are IPHC names that are not given in `iphc-spp-names.csv`.
```{r notconverted}
check_iphc_spp_name()
```

These are the ones just for the new 2020 data:
```{r justnew}
check_iphc_spp_name(countData2020)
```

Of these, only the Thornyhead and Blackspotted Rockfish are likely of interest
(Issues #17 and #18). And maybe the sharks.
