---
title: "Check the 2020 data and incorporate it into gfiphc"
author: "Andrew Edwards"
output: rmarkdown::pdf
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 8
)
```

In 2020 only the first 20 hooks were evaluated, so those data are not going into
GFbio. Going to incorporate into gfiphc here. Likely need this as a template for
future years. This includes some manual checks to make sure the data look okay.

<!-- Using vignette as a template, and iphc2013data.Rnw -->

```{r setup}
load_all()
setData2013
countData2013
```

We want to get the new data into the same format as those. Two data sets are
needed for gfiphc code that to summarises catches of a particular species at the
station level to get counts of zero included.

## Set-level information

Maria was sent the file `2020 IPHCtoDFO_dataExtraction-Maria.xls` for set
details, but this is multiple sheets and more complex than needed. Will try
extracting directly from IPHC website (which they want us to do in the future
anyway), using the following instructions:

Go to https://www.iphc.int/data/fiss-data-query and select the following options:

1. Year Range - 2020 to 2020.

2. Area 2B

3. Purpose Codes - All

4. IPHC Charter Regions - All

5. Maps - Nothing

6. Select non-Pacific halibut species - deselect All.

Download tab on bottom right (see instructions above question 4), and select
CrossTab. Select "Set and Pacific Halibut data" and .xlsx format (I tried .csv
format but it didn't save with commas!?). Save in this
folder as `set-and-halibut-data-2020.xlsx`. Open in Excel and Export as .csv,
and when trying to quit Excel say no to save changes (not sure if that matters).

```{r loadsetdata}
sets_raw <- readr::read_csv("set-and-halibut-data-2020.csv") %>%
  dplyr::mutate_if(is.character, factor)
sets_raw
summary(sets_raw)
testthat::expect_equal(unique(sets_raw$"IPHC Reg Area"), "2B") # Check just BC data
```

Now simplify down to what's needed and rename, based on `iphc2013data.Rnw` (need
to include the 'purpose' column, unlike 2013):
```{r simplify}
sets_simp <- select(sets_raw,
                    year = Year,
                    station = Station,
                    lat = "MidLat fished",
                    lon = "MidLon fished",
                    avgDepthFathoms = "AvgDepth (fm)",
                    hooksRetr = HooksRetrieved,
                    hooksObs = HooksObserved,
                    skatesHauled = SkatesHauled,
                    effSkateIPHC = EffectiveSkates,
                    usable = Effective,
                    purpose = Purpose) %>%
             mutate(E_it20 = effSkateIPHC * hooksObs / hooksRetr,
                    station = as.character(station),
                    usable = as.character(usable))
```

then change purpose to match 2019/2018.


## Species counts














## Data for species of interest

First, just running for multiple individual species.

These are the original species (herring predators) extracted for Joe, though
taking out longnose skate since it will get saved as part of `skate_sp` below.
```{r definesp}
sp_vec <- c("yelloweye rockfish",
            "arrowtooth flounder",
            "lingcod",
            "north pacific spiny dogfish",
            "pacific cod",
            "pacific halibut",
            "redbanded rockfish",
            "sablefish",
            "walleye pollock")
```

We want a general rockfish category and general skates one, I just picked
yelloweye, redbanded, and longnose as examples above. Going through groundfish
synopsis report and picking skate species, then rockfish, that have sufficient
data to look useful.

## Skate species

Unidentified (species of) skates were enumerated in the early years, but then
declined dramatically and had zero counts in 2001, 2002 and 2013, and presumably
for 2003 onwards (but those data are in GFbio which would have to be checked --
seems clear that the identification protocol improved through time). So after
investigating this (in earlier versions of this vignette), am now including
unidentified skate in the species list.

Looking through 2019 synopsis report and the plots below (and a look at the
data), here are the decisions on including each skate species:

 - Aleutian Skate: Looks like only enumerated explicitly since 2007, 2018 map
   shows mean 0.31 fish/skate. Include.
 - Abyssal Skate - no catch, ignore.
 - Broad Skate - no catch, ignore.
 - Big Skate - looks like enumerated explicitly since 1998, and 2018 map shows
   mean 0.69 fish/skate. Include.
 - Roughtail Skate - looks like only caught in three or four years (mean +ve sets
   1/177). No catch in 2018. Include.
 - Sandpaper Skate - only shows up for a few years. Mean +ve sets 5/177, 2018
   mean 0.14 fish/skate. Include.
 - Longnose Skate - only shows up since 1998 (presumably unidentified
   beforehand), mean +ve sets 57/135, 2018 mean 0.96 fish/skate. Include.
 -  Alaska Skate - only showed up in five years, 2018 mean 0.14 fish/skate. Only
   plotted since 2003 (like Sandpaper). Include.

```{r skatedefn}
skate_sp <- c("aleutian skate",
              "big skate",
              "roughtail skate",
              "sandpaper skate",
              "longnose skate",
              "alaska skate",
              "unidentified skate")

skate_ignore <- c("abyssal skate",
                  "broad skate")
skate_unident <- "unidentified skate"
# skates_combined <- "skates combined"
```

Cache the data once (this is not evaluated each time):
```{r get_skates, eval=FALSE}
cache_pbs_data_iphc(skate_sp)
cache_pbs_data_iphc(skate_ignore)
# cache_pbs_data_iphc(skate_unident)
```

Next steps:

 D functionalise next bit, saving results (like in other vignette?)

 D show unident skate also

 D search all years for anything with skate, there are some rare ones not in GFbio

 D add unident and anything else to skate_sp, ensure consistent defns through time

 D create time series for combined skate_sp

 D create time series for the other predators

 - doing a combined rockfish one?

 - Then re-run again with 2020 data added and 2018
    expanded areas accounted for (may be done already for Series EF).

 D create wrapper for format for Jennifer

 - fix Travis

 - merge into master

## Individual skate species

Calculate series for each skate species in turn (saved as `species_name_area`)
and plot figures of Series E and F and longest from those. The fourth panel has
Series EF which is the longest time series possible; if it is blank then Series
E or F are the longest (and automatically identified). But this only happens for
the rare species, and so not when the skate species are combined.

```{r repeat, warning = FALSE, results = FALSE}
for(sp in skate_sp){
  assign(sp_hyphenate(sp, underscore = TRUE, area = TRUE),
         iphc_get_calc_plot_area(sp))
}
```

## Unidentified skate

Above figure shows that for 1995 and 1996 there are 'Unident. skate', and then
for a few years after that (but at lower numbers). Can see from species figures
above that indeed no skate species were identified for those years. And it looks
like then big and longnose were identified but not others for a few years.

First looking into the raw data:

```{r unidentcheck}
# 1995
unique(countData1995$spNameIPHC)[grep("Skate", unique(countData1995$spNameIPHC))]
grep("skate", unique(countData1995$spNameIPHC)) # none
# And a manual check through confirms, that all we need is "unident. Skate"

#1996 to 2002 has "unident. Skate", but only up to 2000:
unique(data1996to2002$spNameIPHC)[grep("Skate",
                                       unique(data1996to2002$spNameIPHC))]
grep("skate", unique(data1996to2002$spNameIPHC)) # none
dplyr::filter(data1996to2002,
              spNameIPHC == "unident. Skate") %>%
  tail()

# And there were less unidentified through time:
dplyr::filter(data1996to2002,
              spNameIPHC == "unident. Skate") %>%
  select(year) %>%
  table()


# 2013 (not in GFbio), no unident. Skate:
unique(countData2013$spNameIPHC)[grep("Skate", unique(countData2013$spNameIPHC))]
grep("skate", unique(countData2013$spNameIPHC)) # none

# 2003-2019 (except 2013) - in GFbio
# See Issue #16 - not sure what the species code would be, but there may not be
# one if they were all identified after 2000.

# 2020 (not in GFbio). Not looked at yet. Presumably identified. TODO
```

So, given none in 2001 or 2002 and numbers of unidentified declined from 1996 to
2000, it's probably the case that there were all identified from 2001
onwards. See Issue #16.

So, we can use `unidentified skate` as a species name (done above). Load in and
examine that data:
```{r unidentsummary}
unidentified_skate_area$sp_set_counts_with_area %>%
  select(year) %>%
  table()
```
But there no positive counts after 2000, looking at first 20 hooks (Series E)
and then all hooks (Series F).
```{r unidentsummary2}
dplyr::select(unidentified_skate_area$ser_E_and_F$ser_E,
              year,
              num_pos20)
dplyr::select(unidentified_skate_area$ser_E_and_F$ser_F,
              year,
              num_pos)
```

## Merge together skate species

Now want to lump the skates in `skate_sp` together. Note that species counts get
named as `N_it.1` etc. While each one has NA's, none of the summed `N_it_sum` or
`N_it20_sum` do, because `get_combined_species()` ignores NA's when making the
sums.

```{r combine, warning = FALSE}
sp <- "skates combined"
skates_combined <- get_combined_species(sp_vec = skate_sp,
                                        save_RDS_name = sp_hyphenate(sp))
summary(skates_combined$set_counts)

# How many NA's in certain columns (none for N_it_sum or N_it20_sum
skates_combined$set_counts %>%
  dplyr::select(-c("station", "lat", "lon", "usable", "standard")) %>%
  dplyr::summarise_all( ~ sum(is.na(.x))) %>%
  t()
```

Now use to calculate Series E, F and EF calculations for the restricted area:
```{r combineEF, warning = FALSE, results = FALSE}
assign(sp_hyphenate(sp, underscore = TRUE, area = TRUE),
       iphc_get_calc_plot_area(sp))
```

## Individual non-skate species

Similarly for the other species:
```{r showmult, warning = FALSE, results = FALSE}
for(sp in sp_vec){
  assign(sp_hyphenate(sp, underscore = TRUE, area = TRUE),
         iphc_get_calc_plot_area(sp))
}
```

## Format for Jennifer's EAFM summary

Save annual indices (with uncertainty) for all species of interest, in the
format for the HG herring predators' EAFM summary.

```{r allsp}
all_sp_longest <- list()
sp_vec_and_combined_skates <- c(sp_vec, "skates combined")
# Do this to easily create arguments for iphc_format_for_EAFM(); can't easily
#  functionalise. Each element is names for the species, and is a tibble of
#  longest time series.
for(sp in sp_vec_and_combined_skates){
  all_sp_longest[[sp]] <- get(paste0(sp_hyphenate(sp,
                                                  underscore = TRUE,
                                                  area = TRUE)))$series_longest$ser_longest
}

eafm_res <- iphc_format_for_EAFM(sp_vec_and_combined_skates,
                                 all_sp_longest) # also saves .csv, and .rds of
                                        # all_sp_longest list

eafm_res
```

Note that the code may not yet fully account for all possibilities (for all
areas and species), particularly for rarer species. So do check that the output
is sensible.
