---
title: "Check the 2021 data and incorporate it into gfiphc"
author: "Andrew Edwards"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ">",
  fig.width = 10,
  fig.height = 8
)
```

In 2021 (as for 2020) only the first 20 hooks were evaluated, so those data are not easily imported into
GFBio. Going to incorporate into gfiphc here. Likely need this as a template for
future years: resave this file with new year, and change all 2021's to the
subsequent year, go through the code somewhat manually to check the output
as you go along (in Emacs do Alt-query-replace to change years but read
carefully as going along), and then finally render the full document to make the
.pdf. This code includes some
manual checks to make sure the data look
okay. The planned stations for the 2021 survey are shown in this [IPHC
sampling manual (click here)](https://www.iphc.int/uploads/pdf/fiss/2021/iphc-2021-fiss-stations.pdf);
page 23 has Vancouver [Island] Outside, showing that not all stations were
intended to be fished there.

For comparison first look at 2013 data included in gfiphc:
<!-- Based on iphc2013data.Rnw -->
```{r setup}
load_all()
setData2013
countData2013
```

We want to get the new data into the same format as those (columns with same
names and classes, even though in retrospect some classes aren't ideally
chosen, but also retaining retrieved and observed hooks for the set data). Two
data sets are
needed because later gfiphc code summarises catches of a particular species at the
station level, and needs to create counts of zeros for the species of interest
(and such zeros are not included in IPHC output).

## Set-level information

For 2020, Maria was sent the file `2020 IPHCtoDFO_dataExtraction-Maria.xls` for set
details, but this is multiple sheets and more complex than needed. So I tried
extracting directly from the IPHC website (which they want us to do in the future
anyway), using the following instructions, which worked for 2020 and 2021:

Go to https://www.iphc.int/data/fiss-data-query and select the following options:

1. Year Range -- 2021 to 2021.

2. Area 2B

3. Purpose Codes -- All

4. IPHC Charter Regions -- All

5. Maps -- Nothing

6. Select non-Pacific halibut species -- deselect All.

Download tab on bottom right (see instructions above question 4), and select
CrossTab. Select "Set and Pacific Halibut data" and .xlsx format (I tried .csv
format but it didn't save with commas, strangely). Save in this
folder as `set-and-halibut-data-2021.xlsx`. Open in Excel and Export as .csv, `set-and-halibut-data-2021.csv`,
and when trying to quit Excel say no to save changes (not sure if that matters).

Repeat but with all non-halibut data (select All in number 6), and save as
`non-halibut-data-2021.xlsx` and export as .csv in Excel,
`non-halibut-data-2021.csv`. Importantly, this file
(but not the first one) contains the numbers of observed hooks, needed in our
calculations.

Load data for new year:
```{r loadsetdata}
sets_raw <- readr::read_csv("set-and-halibut-data-2021.csv") %>%
  dplyr::mutate_if(is.character, factor)
```

Now load the original 2020 data (do not change the 2020 here) to then test that the column
names and types do not change in future years, and then check columns match `sets_raw`:
```{r load2020setdata}
sets_raw_2020 <- readr::read_csv("set-and-halibut-data-2020.csv") %>%
  dplyr::mutate_if(is.character, factor)

# For 2021 these were different - uncomment for future for first test
# testthat::expect_equal(names(sets_raw_2020),
#                        names(sets_raw))

# testthat::expect_equal(sapply(sets_raw_2020, typeof),
#                        sapply(sets_raw, typeof))

# Columns in 2020 not in new data:
setdiff(names(sets_raw_2020), names(sets_raw))

# Columns in new data not in 2020:
setdiff(names(sets_raw), names(sets_raw_2020))

# For 2021 looks like Purpose became Purpose Code, but are the same type:
summary(sets_raw_2020$Purpose)
summary(sets_raw$"Purpose Code")
testthat::expect_equal(typeof(sets_raw_2020$Purpose),
                       typeof(sets_raw$"Purpose Code"))
```

Those extra columns in 2021 look related to oceanographic data, beyond the scope
of gfiphc, so can just ignore shortly.

Want to check the overlapping columns have the same type:
```{r overlappingcols}
overlap_col_names <- intersect(names(sets_raw_2020),
                               names(sets_raw))
# testthat::expect_equal(sapply(dplyr::select(sets_raw_2020,
#                                            overlap_col_names),
#                              typeof),
#                       sapply(dplyr::select(sets_raw,
#                                            overlap_col_names),
#                              typeof))
# Error: sapply(dplyr::select(sets_raw_2020, overlap_col_names), typeof) not equal to sapply(dplyr::select(sets_raw, overlap_col_names), typeof).
# 1/32 mismatches
# x[12]: "logical"
# y[12]: "integer"
dplyr::select(sets_raw_2020,
              overlap_col_names[12])
dplyr::select(sets_raw,
              overlap_col_names[12])
```
These are all NA's anyway (see below) and don't get saved, so no worries.


```{r newsetdata}
sets_raw
summary(sets_raw)
testthat::expect_equal(unique(sets_raw$"IPHC Reg Area"),
                       as.factor("2B")) # Check just BC
testthat::expect_equal(unique(sets_raw$Year), 2021)
testthat::expect_equal(length(unique(sets_raw$Station)),
                       length(sets_raw$Station))
```

### Understand any issues raised above

Uncomment those three `testthat` commands when looking at new data each year. If
any of fail then have to comment it out and figure out what it means here.

This is for 2020 (check for future years), to look for station(s) that was
fished twice. Not really needed for 2021 since that third test passed, but
`twice_fished` gets used later, so do evaluate here:
```{r understand}
length(unique(sets_raw$Station))
length(sets_raw$Station)
dplyr::count(sets_raw, Station) %>% dplyr::filter(n > 1)
twice_fished <- dplyr::count(sets_raw, Station) %>%
  dplyr::filter(n > 1) %>%
  dplyr::select(Station) %>%
  as.numeric()
twice_fished
# If there's more than a single station then adapt later code
#as.data.frame(dplyr::filter(sets_raw,
#                            Station == twice_fished))

```
Not needed for 2021: So Station `r twice_fished` had two vessels fishing the same station (which the code below
originally caused a total of four rows for that station, explaining the 200 rows
I had in original `setData2020` before fixing the issue). Interestingly the
halibut catches were almost double for one vessel than the other (but were 6
days apart):

2020: Note that one of those entries has 'Vessel code' HAN, but HAN only appears once
in the whole data set (as seen in `summary(sets_raw)` above.

For 2021, just noting that two vessels were used, and these are different to those in 2020 (for which HAN then
got excluded anyway):
```{r duplicate}
summary(sets_raw$"Vessel code")
summary(sets_raw_2020$"Vessel code")
```

2020: So given we want to exclude one of the duplicates, makes sense to exclude
HAN. (Also, Dana mentioned some gear comparison studies for 2020).

Simplify down to what's needed and rename, based on `iphc2013data.Rnw` (need
to include the 'purpose' column, unlike 2013):
```{r simplify}
# sets_simp <- dplyr::filter(sets_raw, `Vessel code` != "HAN") %>%
sets_simp <- dplyr::select(sets_raw,
                           year = Year,
                           station = Station,
                           lat = "MidLat fished",
                           lon = "MidLon fished",
                           avgDepth = "AvgDepth (fm)",
                           skatesHauled = "No. skates hauled",
                           effSkateIPHC = "Effective skates hauled",
                           soakTimeMinutes = "Soak time (min.)",  # Joe might want
                           usable = Eff,
                           purpose = "Purpose Code",
                           U32halibut = "U32 Pacific halibut count",
                           O32halibut = "O32 Pacific halibut count") %>%
  arrange(station) %>%
  dplyr::mutate(year = as.integer(year),
                station = as.character(station),
                avgDepth = as.integer(avgDepth),
                usable = as.character(usable))
sets_simp
```

## Standard grid or not

Need to change purpose to `standard` (`Y/N`) to match 2018 data (`Y` for the
standard grid). In the raw 2020 data, `Purpose` took three values that we
converted to `standard` to save in the package:
```{r purpose2020}
summary(sets_raw_2020$Purpose)
summary(setData2020$standard)
```

For 2021 we have all as Standard Grid, which gets corrected (some stations are
non-standard) in the next section.
```{r purpose}
summary(sets_simp$purpose)

sets_simp_std <- dplyr::mutate(sets_simp,
                               standard_tmp = (purpose == "Standard Grid"))
                               # was grid in 2020, Grid in 2021

standard <- as.character(sets_simp_std$standard_tmp)  # to get the right length
standard[sets_simp_std$standard_tmp] = "Y"
standard[!sets_simp_std$standard_tmp] = "N"
length(standard)

sets_simp_std <- cbind(sets_simp_std,
                   standard) %>%
  as_tibble() %>%
  dplyr::select(-c("standard_tmp"))
summary(sets_simp_std)
unique(sets_simp_std$standard)
```

So they are all classified as standard. For 2020 we stuck with the 2018
definitions of standard, so doing that next.

## Look at data and show map to understand changing definition of standard station from 2018 to 2020.

The definition of 'standard grid' changed from 2018 (when
first needed due to the expanded grid) to 2020 (and 2021). Simply equating them as above is
not sufficient. For 2021 we so far have this:
```{r map2021}
plot_iphc_map(sets_simp_std,
              sp = NULL,
              years = 2021,
              indicate_standard = TRUE)
```
So no stations are marked as being outside the standard grid, even though some
are clearly new -- the ones in the north have never been fished before (see the
one-species vignette, though I'll investigate that here).

This next section was to first figure out the twice-fished station `2343` in 2020, and to
replicate that original analysis (station ends up being non-standard later), so
mostly commented out except first bit which is used later so keeping in case
need in future years:
```{r revert}
hooks_with_bait_revert <- hooks_with_bait

# This should be commented out for 2021 survey analysis in iphc-2021-data.Rmd,
#  since the problem is presumably fixed. This is to revert back to the original
#  problem, for which 2343 was called standard in 2018 but we changed it. Map on
#  page 10 of iphc-2020-data.pdf has this station (second one down off
#  north-east tip of Haida Gwaii) as non-standard in 2018 but not 2020.
# hooks_with_bait_revert$set_counts[hooks_with_bait_revert$set_counts$year == 2018 &
#                            hooks_with_bait_revert$set_counts$station == 2343,
#                            ]$standard = "Y"

#filter(hooks_with_bait$set_counts, year == 2018, station == 2343) %>%
#  as.data.frame()      # saved version
#filter(hooks_with_bait_revert$set_counts, year == 2018, station == 2343) %>%
#  as.data.frame()      # reverted version

```

Now to figure out standard/non-standard stations. Plotting four years, with crosses showing 'non-standard'. (2021 is coloured
different since no hooks with bait data yet, but the important bit is the crosses).
```{r maps}
sets_2021 <- dplyr::select(sets_simp_std,
                           -c(U32halibut, O32halibut))
                           # else not the same structure as sets_2018, below

plot_iphc_map_panel(hooks_with_bait$set_counts,
                    sp = "Hooks with bait",
                    years = 2018:2020,
                    indicate_standard = TRUE)


plot_iphc_map(sets_2021,
              sp = NULL,
              years = 2021,
              indicate_standard = TRUE)
```

Can see that 2020 has a few less stations just north of Vancouver Island, but not
enough to worry about greatly, and 2021 has kind of done a few of those. The
2021 ones way in in the inlets are not currently flagged as non-standard but
will be below (using the 2018 definitions). In fact no stations are flagged for
2021 as non-standard. And the other main issue is that 2021 is doing a random
sample of WCVI stations (some of which will become non-standard). AND that there
are new stations in the north (and maybe elsewhere) that have never been fished
before (as I discovered when updating the one-species vignette and redefining
the default axes limits for `plot_BC()`; the version before updating that isaved
as iphc-2021-data-all-2021-stations.pdf`). Will examine those shortly.

Need to look and plot values:
```{r standardornot}
sets_2018 <- filter(hooks_with_bait_revert$set_counts,
                    year == 2018)
not_std_2018 <- filter(sets_2018,
                       standard == "N")$station

not_std_2021 <- filter(sets_2021,
                       standard == "N")$station

# Not standard in both:
not_std_2018_and_2021 <- intersect(not_std_2018, not_std_2021)
not_std_2018_and_2021

length(not_std_2018)
length(not_std_2021)
length(not_std_2018_and_2021)

# 2018 has some east of the map, all non-standard:
filter(hooks_with_bait_revert$set_counts, year == 2018, lon > -124)$standard
nrow(filter(hooks_with_bait_revert$set_counts, year == 2018, lon > -124))

std_in_2018_but_not_std_in_2021 <- intersect(filter(sets_2018,
                                                    standard == "Y")$station,
                                             not_std_2021)
std_in_2018_but_not_std_in_2021

not_std_in_2018_but_std_in_2021 <- intersect(not_std_2018,
                                             filter(sets_2021,
                                                    standard == "Y")$station)
not_std_in_2018_but_std_in_2021

# setdiff(x, y) - elements in x but not in y
# setdiff(not_std_2018, not_std_2020) - but 2020 fewer coverage so misleading
```

Plot stations not standard in 2018 but standard in 2021, and vice versa, using
each years' lats and lons (to verify that they all still agree -- i.e., that station
numbers have consistent lats and lons), and show 2019 data to check no 'usual'
stations are non-standard in 2018 or 2021. Also (for 2021) adding all stations,
since this will clearly show the random sampling off WCVI:
```{r mapstandard, out.width = '120%', fig.align = "center"}
plot_BC()
points(lat~lon,
       data = filter(sets_2018,
                     station %in% not_std_in_2018_but_std_in_2021),
       col="red",
       pch = 19)

# Do the same but using 2021 station co-ordinates - should overlap:
points(lat~lon,
       data = filter(sets_2021,
                     station %in% not_std_in_2018_but_std_in_2021),
       col="blue",
       pch = 3)

# And for 2020 showed the single station std in 2018 but not 2020, for 2021
# there are none:
points(lat~lon,
       data = filter(sets_2018,
                     station %in% std_in_2018_but_not_std_in_2021),
       col="red",
       pch = 17)
points(lat~lon,
       data = filter(sets_2021,
                     station %in% std_in_2018_but_not_std_in_2021),
       col="blue",
       pch = 1,
       cex = 2)


# Now show all 2019 stations:
points(lat~lon,
       data = filter(hooks_with_bait_revert$set_counts,
                     year == 2019),
       col="darkgreen",
       pch = 0)

# Add all 2021 stations as a small black dot
points(lat~lon,
       data = sets_2021,
       col="black",
       pch = 20,
       cex = 0.8)

legend("bottomleft",
       legend = c("Not std 2018 but std 2021 (2018 co-ords)",
                  "Not std 2018 but std 2021 (2021 co-ords)",
                  "Std 2018 but not std 2021 (2018 co-ords)",
                  "Std 2018 but not std 2021 (2021 co-ords)",
                  "All 2019 stations",
                  "All 2021 stations"),
             pch = c(19, 3, 17, 1, 0, 20),
             pt.cex = c(1, 1, 1, 2, 1, 0.8),
             col = c("red", "blue", "red", "blue", "darkgreen", "black"))
```
So the co-ordinates look close enough (red circles and blue crosses overlap),
none were defined as non-standard in 2021 so there are no red triangles or blue
circles, and the green squares for 2019 stations correctly do not overlap with
the non-standard 2018 stations. Black dots (2021 stations) with no green squares
off WCVI clearly shows the reduced coverage there.

2020 only (there were no non-standard stations defined in raw data for 2021):
Check if the one standard station in 2018 but not in 2020 (not fished at all in 2019)
appears in any earlier years:
```{r checkearly, eval=FALSE}
# Fails (so not evaluated here) since empty in 2021, and this is corrected
dplyr::filter(hooks_with_bait_revert$set_counts,
              station == std_in_2018_but_not_std_in_2021) %>%
  as.data.frame()
```
For 2020 I worked out it was only fished in 2018 and 2020 so we defined it as non-standard.

So, the conclusion from this section so far is that we should retain the 2018
definitions of standard stations, not the new ones defined in
2021, as we did for 2020.

Doing that shortly (in `sets_simp_std_corrected`), but first also look for any
new 2021 stations. I hadn't expected any but saw them when doing the one-species
vignette, so had to come back to redo this.

```{r stations2021only}
# yelloweye_rockfish$set_counts is saved in gfiphc, already has 2021 data
#  because I had to come back to redo this .pdf after updating the data, hence
#  need the <2021 here; station codes do change over time, but I think are
#  recently consistent
previous_stations <- dplyr::filter(yelloweye_rockfish$set_counts,
                                   year < 2021)$station %>%
                     unique()
stations_in_2021_only <- dplyr::filter(sets_2021,
                                       !(station %in% previous_stations))
stations_in_2021_only
```
and plot those stations:
```{r map2021only, out.width = '120%', fig.align = "center"}
plot_BC()
points(lat~lon,
       data = stations_in_2021_only,
       col = "black",
       pch = 15)

points(lat~lon,
       data = dplyr::filter(yelloweye_rockfish$set_counts,
                            year < 2021),
       col = "red",
       pch = 20,
       cex = 0.4)

legend("bottomleft",
       legend = c("Station numbers appearing only in 2021 and never before",
                  "Stations from all previous years"),
       pch = c(15, 20),
       col = c("black", "red"),
       pt.cex = c(1, 0.4))
```

So there are six 2021 stations that have never been fished before! That map
suggests that we should call the five northern ones non-standard also, to exclude
from the standard Series A-F analyses.

However, Ann-Marie Huang thinks that
these stations may have been fished before but considered as part of Area 2C
(Alaskan waters). Some waters around there are claimed by both Canada and the
US; there's a clear map and explanation in [Canada's Unresolved Maritime
Boundaries
(clickable)](https://www.dur.ac.uk/resources/ibru/publications/full/bsb5-3_gray.pdf),
which is linked from [this Wikipedia article on Dixon
Entrance](https://en.wikipedia.org/wiki/Dixon_Entrance).
So there may be earlier data, which are not in gfiphc because such stations
would not have been considered Area 2B, which is the area for which the IPHC
sent DFO data in the past (and which I used here for recent years to extract
from their website). So there may be data available, and if needed it will have
to be obtained. Here we will call those five northern newly-fished stations
`non-standard`.

For the sixth station off the northwest of Vancouer Island, zooming in and
including the Scott Islands [Rockfish Conservation Area
(clickable)](https://www.pac.dfo-mpo.gc.ca/fm-gp/maps-cartes/rca-acs/rca-acs/north-nord/ScottIslandsChart3744-eng.html)
as a blue rectangle shows:
```{r map2021onlyzoom, out.width = '120%', fig.align = "center"}
plot_BC(xlim = c(-130, -127),
        ylim = c(50, 52))

scott_island_RCA_lon <- -c(128 + 56.5/60, 128 + 33/60)
scott_island_RCA_lat <- c(50 + 45/60, 50 + 52/60)

# rect(xleft, ybottom, xright, ytop, density = NULL, angle = 45,
rect(scott_island_RCA_lon[1],
     scott_island_RCA_lat[1],
     scott_island_RCA_lon[2],
     scott_island_RCA_lat[2],
     border = "blue")

points(lat~lon,
       data = stations_in_2021_only,
       col = "black",
       pch = 15)

points(lat~lon,
       data = dplyr::filter(yelloweye_rockfish$set_counts,
                            year < 2021),
       col = "red",
       pch = 20,
       cex = 0.4)

legend("bottomleft",
       legend = c("Station numbers appearing only in 2021 and never before",
                  "Stations from all previous years"),
       pch = c(15, 20),
       col = c("black", "red"),
       pt.cex = c(1, 0.4))
```
So the new station is just outside the RCA. Presumably in previous years the RCA
was avoided as the grid would have put a station in the RCA, close to (or even
on) Lanz Island.

It is station 2257 (see above), with a depth of only
40 fathoms, which is not an outlier. For example, for 2013
(depth data for all years is not in gfiphc I don't think):
```{r depth2013}
sort(setData2013$avgDepth)
```
However, since it is a new station and not been used before, we will flag it as
`non-standard` (as used for the Series A-F analyses). Also Dana Haggarty says
that there is good habitat right close to those islands, but not great further
away, and she has used Remotely Operated Vehicles there -- it's all
sand/gravel/cobble with massive sand waves from the crazy exposure, but perhaps
there are pockets of good habitat. So either
way (close to an RCA so may be expected to be good for rockfish at least, or not
great rockfish habitat) it shouldn't really be included for rockfish species,
and in general should be excluded since a new station.

So - retain the 2018 definitions of standard stations (as we did for 2020),
and call all six 2021 stations non-standard:
```{r corrected}
sets_simp_std_corrected <- sets_simp_std
summary(as.factor(sets_simp_std_corrected$standard))

sets_simp_std_corrected$standard[sets_simp_std_corrected$station %in%
                                 not_std_in_2018_but_std_in_2021] <- "N"
sets_simp_std_corrected$standard[sets_simp_std_corrected$station %in%
                                 stations_in_2021_only$station] <- "N"
summary(as.factor(sets_simp_std_corrected$standard))
# cbind(sets_simp_std$standard, sets_simp_std_corrected$standard) # to check them
```

Think I hadn't originally defined them as factors in early code, so keeping them
as characters now. Just to vefify that none of the 2018 non-standard stations
were fished before 2018:
```{r checknon}
dplyr::filter(hooks_with_bait$set_counts,
              station %in% not_std_2018) %>%
  dplyr::select(year) %>%
  unique()
```
Note 2021 won't show up here until .rda objects are resaved in package, at the
end of this .pdf (so it will if this .Rmd has already been run, as it has in 2021).

So here are the final station designations for 2021:
```{r map2021final}
plot_iphc_map(sets_simp_std_corrected,
              sp = NULL,
              years = 2021,
              indicate_standard = TRUE)
```
Can see that they did 10 random stations off WCVI that we're calling
non-standard (because they were never fished before 2018). Which is a bit of a
shame as there are only 16 stations left off WCVI for 2021.

2020 (no need to change for 2021): So check which functions need changing, since they create a 'standard'
column. These do not need changing: `get_iphc_hooks()` and `get_iphc_skates_info`.

2020: Then `get_iphc_sets_info()` does return `standard`, but the `standard`
designation is not saved in GFBio it is saved in `setDataExpansion` in `gfiphc`.
So just need to add a line in `IPHC-stations-expanded.R` and then re-save all
.rda files. Fixed that, now recreating all .rda files, as per the README.

## Species counts

Now get the species counts into the desired format (to match `countData2013`
shown earlier). First check that the column names and types haven't changed
(they did for set data from 2020 to 2021):
```{r countsnames}
counts_raw_2020 <- readr::read_csv("non-halibut-data-2020.csv") %>%
  dplyr::mutate_if(is.character, factor)

counts_raw <- readr::read_csv("non-halibut-data-2021.csv") %>%
  dplyr::mutate_if(is.character, factor)

testthat::expect_equal(names(counts_raw_2020),
                       names(counts_raw))

testthat::expect_equal(sapply(counts_raw_2020, typeof),
                       sapply(counts_raw, typeof))
```
Great, nothing changed in the structure.

```{r counts}
counts_raw
summary(counts_raw)
testthat::expect_equal(unique(counts_raw$Year), 2021)  # All 2021
testthat::expect_equal(unique(counts_raw$SampleType), as.factor("20Hook")) # All 20Hook

# This mismatches for 2020, not for 2021:
testthat::expect_equal(length(unique(counts_raw$Station)),
                       length(sets_raw$Station))

unique(counts_raw$"Species Name")
```

Here's what was seen in 2020 but not 2021, and vice versa:
```{r seenoneyear}
# Seen in 2020 not 2021
setdiff(unique(counts_raw_2020$"Species Name"),
        unique(counts_raw$"Species Name"))
# Seen in 2021 not 2020
setdiff(unique(counts_raw$"Species Name"),
        unique(counts_raw_2020$"Species Name"))
```
Presumably Sun Sea Star and Sunflower Sea Star are the same. Will mention this later on.

Note that halibut are not included in these counts:
```{r nohalibut}
dplyr::filter(counts_raw, "Species Name" == "Pacific Halibut")
```
which I presume explains why
total number of counts for a station does not add up to `HooksObserved`. See
later for halibut calculations.

2020 only: Need to remove the HAN records for the twice-fished station, which turns out to
be set number 4 for station 2104:
```{r twicefished}
dplyr::filter(counts_raw, Station == twice_fished) %>%
  dplyr::select(c("Station", "Setno", "Species Name",
                  "Number Observed")) %>%
    as.data.frame()

dplyr::filter(sets_raw, Station == twice_fished)
```

So for 2020 had to use that here to remove the species counts for that vessel (note that vessel
code is not in `counts_raw`), just commenting that part out for 2021:
```{r countcalcs}
dplyr::filter(counts_raw,
              Station == twice_fished & Setno == 4)

# So just keep these:
# dplyr::filter(counts_raw,
#              !(Station == twice_fished & Setno == 4))

#countData2020_no_halibut <- dplyr::filter(counts_raw,
#                               !(Station == twice_fished & Setno == 4)) %>%
# Seems that can't just keep using that even if twice_fished = NA
countData2021_no_halibut <- counts_raw %>%
  dplyr::select(year = Year,
                station = Station,
                spNameIPHC = "Species Name",
                specCount = "Number Observed") %>%
  arrange(station) %>%
  dplyr::mutate(year = as.integer(year),
                station = as.character(station),
                spNameIPHC = as.character(spNameIPHC),
                specCount = as.integer(specCount))

testthat::expect_equal(names(countData2013), names(countData2021_no_halibut))
countData2021_no_halibut
summary(countData2021_no_halibut)
```

## Hooks observed and retrieved

Now, obtain the numbers of hooks observed and retrieved from `counts_raw`, to
then merge into the set details:
```{r hookdetails}
# hook_details <- dplyr::filter(counts_raw,
#                               !(Station == twice_fished & Setno == 4)) %>%
hook_details <- counts_raw %>%
  dplyr::group_by(Station) %>%
  dplyr::summarise(year = unique(Year),
                   hooksRetr = unique(HooksRetrieved),
                   hooksObs = unique(HooksObserved)) %>%
  dplyr::rename(station = Station) %>%
  dplyr::ungroup() %>%
  arrange(station) %>%
  dplyr::mutate(year = as.integer(year),
                station = as.character(station))

hook_details

testthat::expect_equal(sets_simp_std_corrected$station, hook_details$station)
```

So now need to get the hook details into the set details, and keep columns as
for `setData2013` but also with `standard`, and may as well keep `hooksRetr` and
`hooksObs`:
```{r merge}
setData2021 <- dplyr::left_join(sets_simp_std_corrected,
                                hook_details,
                                by = c("year", "station")) %>%
  dplyr::mutate(E_it20 = effSkateIPHC * hooksObs / hooksRetr) %>%
  dplyr::select(year,
                station,
                lat,
                lon,
                avgDepth,
                effSkateIPHC,
                E_it20,
                usable,
                standard,
                hooksRetr,
                hooksObs) %>%
  dplyr::mutate(year = as.integer(year),
                station = as.character(station),
                avgDepth = as.integer(avgDepth),
                usable = as.character(usable),
                standard = as.factor(standard))
setData2021
testthat::expect_equal(names(setData2013), names(setData2021)[1:ncol(setData2013)])
summary(setData2021)
```

## Pacific Halibut counts

As noted above, the data extraction for the counts is for all non-halibut
species. We still want the halibut counts for just the first 20 hooks -- the
`data_for_all_species` vignette (for data up to 2019) shows that the 20-hook and full hook counts
(Series A and B) are very similar when rescaled, and the rescaling is miniscule
with $G_A / G_B = 1.005$. So this justifies sticking with 20-hook counts for
halibut, even though the full data are available for all sets, given it is a
halibut survey. (Using all hooks for all years could be done, but would be a lot
of new code).

There are two options for getting halibut counts for the first 20 hooks (given
we don't have hook-by-hook data, though it could probably be obtained just maybe
not from the IPHC website).

### Option 1.

Take the halibut counts for all the hooks (which we have in `sets_raw` and
subsequent objects) and create `N_it20_halibut_est = E_it20 / E_it *
N_it`, or equivalently just `N_it20_halibut_est = hooksObs / hooksRetr *
N_it`. Note that observed refers to observed for non-halibut species
(presumably `hooksRetr` works for halibut). Not
strictly the first 20 hooks, but is a rescaling. But will not guarantee integer values.

```{r sethalibut}
setData2021_and_halibut <-
  dplyr::left_join(setData2021,
                   dplyr::select(sets_simp_std_corrected,
                                 c(station,
                                   U32halibut,
                                   O32halibut)),
                   by = "station") %>%
  dplyr::mutate(N_it_halibut = U32halibut + O32halibut,
                N_it20_halibut_opt_1 = hooksObs / hooksRetr * N_it_halibut)
setData2021_and_halibut %>% dplyr::select(station,
                                          N_it_halibut,
                                          N_it20_halibut_opt_1)
```

### Option 2.

Add all the 20-hook counts for a set (which include `Hook with Skin` etc.) and compare
   with `hooksObs`. The latter is higher (or equal), and the difference is halibut
   (as the only **non** non-halibut` species). Compare with the results from
   option 1. If close then use option 2, since it will be just be halibut counts
   and gives an integer number, and is based on the first 20 hooks.

Add counts for each set:
```{r countsmaybe}
counts_20 <- countData2021_no_halibut %>%
  dplyr::group_by(station) %>%
  dplyr::summarise(non_halibut = sum(specCount)) %>%
  dplyr::ungroup()
counts_20
```

Now join the two options together to calculate `N_it20_halibut_opt_2` and then
compare the two estimates of `N_it20_halibut`:
```{r compare}
compare <-
  dplyr::left_join(setData2021_and_halibut,
                   counts_20,
                   by = "station") %>%
  dplyr::mutate(N_it20_halibut_opt_2 = hooksObs - non_halibut,
                N_it20_opt_1_over_opt_2 = N_it20_halibut_opt_1 / N_it20_halibut_opt_2) %>%
  dplyr::select(year,
                station,
                usable,
                N_it20_halibut_opt_1,
                N_it20_halibut_opt_2,
                N_it20_opt_1_over_opt_2)
compare$spNameIPHC <- "Pacific Halibut"
compare

plot(compare$N_it20_halibut_opt_1, compare$N_it20_halibut_opt_2)
abline(a = 0, b = 1)

cor(compare$N_it20_halibut_opt_1,
    compare$N_it20_halibut_opt_2)
```
So this is the right approach and correlation coefficient is high, though
numbers not quite as close as may have thought. But these data are used for
aggregating across all stations in a year
(and any further analyses on halibut for management purposes should be done
using the full halibut data
anyway -- we wouldn't really need that). And the means aren't too bad:
```{r means}
mean(compare$N_it20_halibut_opt_1)
mean(compare$N_it20_halibut_opt_2)
```

So either of these would work. So use option 2 since gives an integer count:
```{r combine}
compare$N_it20_halibut_opt_2
countData2021_halibut <- dplyr::select(compare,
                                       year,
                                       station,
                                       spNameIPHC,
                                       specCount = N_it20_halibut_opt_2) %>%
  dplyr::mutate(specCount = as.integer(specCount))
countData2021 <- rbind(countData2021_no_halibut,
                       countData2021_halibut) %>%
  dplyr::arrange(station)
# First time running, called the above countData2020_NEW to check remaining data didn't change:
# expect_equal(countData2020, filter(countData2020_NEW, spNameIPHC !=
#                                                      "Pacific Halibut"))
```

Note that for 2021 this does give zeros for Pacific Halibut (the only species that
will have a zero, because we have a value for each station because zero counts are in
the original `sets_raw`):
```{r countData}
summary(dplyr::filter(countData2021,
                      spNameIPHC == "Pacific Halibut"))
unique(dplyr::filter(countData2021, specCount == 0)$spNameIPHC)
```

## Check species names

The file `inst/extdata/iphc-spp-names.csv` contains species common names (as
used for gfsynopsis, and a few extra like `unidentified skate`) and the IPHC
common name. The function `check_iphc_spp_name()` has a list of non-groundfish
species that are automatically ignored. These first results are from running
these functions *before* updating anything, so the results are hardwired
here (chunks are not evaluated). Then we update the species list and re-run the functions.

These are IPHC names that are not given in `iphc-spp-names.csv` (automatically ignoring
obvious ones that are listed in the function), for years up to 2020 (since not
updated code yet):
```{r notconverted, eval=FALSE}
check_iphc_spp_name()
##  [1] "Unidentified Shark"          "Unident. Rockfish"
##  [3] "unident. thornyhead (Idiot)" "Grenadier (Rattails)"
##  [5] "Miscellaneous Shark"         "Eelpout"
##  [7] "unident. Roundfish"          "unident. Sculpin"
##  [9] "Unident. Flatfish"           "Greenland Turbot"
## [11] "unident. Hagfish"            "Starry Skate"
## [13] "Black Skate"                 "Brittle Star"
## [15] "Glass Sponge"                "Basketstar"
## [17] "Blackspotted Rockfish"
```

These are the ones just for the new 2021 data:
```{r justnew, eval=FALSE}
check_iphc_spp_name(countData2021)
##  [1] "Basketstar"                   "unident. thornyhead (Idiot)"
##  [3] "Sandpaper Skate"              "Sea Whip"
##  [5] "Stylaster campylecus (coral)" "Brittle Star"
##  [7] "unident. Sculpin"             "Glass Sponge"
##  [9] "Sun Sea Star"                 "Salmon Shark"
## [11] "Jellyfish"                    "Great Sculpin"
## [13] "Cabezon"                      "Unident. Salmon"
## [15] "Unident. Rockfish"            "unident. organic matter"
## [17] "Dungeness Crab"               "Blackspotted Rockfish"
```
There were only six for 2020 though (a lot more for 2021):
```{r new2020, eval=FALSE}
check_iphc_spp_name(countData2020)
## [1] "unident. thornyhead (Idiot)" "Brittle Star"
## [3] "Glass Sponge"                "Basketstar"
## [5] "Blackspotted Rockfish"       "unident. Sculpin"
```

For 2020 I said that only the Thornyhead and Blackspotted Rockfish are likely of
interest (Issues #17 and #18). And the sharks from the earlir list. So look at just the new ones in 2021 that
aren't in 2020 or any previous year:

```{r new2021, eval =FALSE}
setdiff(check_iphc_spp_name(countData2021),
        check_iphc_spp_name())
# Before updating anything this gives:
# [1] "Sandpaper Skate"              "Sea Whip"
# [3] "Stylaster campylecus (coral)" "Sun Sea Star"
# [5] "Salmon Shark"                 "Jellyfish"
# [7] "Great Sculpin"                "Cabezon"
# [9] "Unident. Salmon"              "unident. organic matter"
# [11] "Dungeness Crab"
```

Of these, Sandpaper Skate, Salmon Shark, Great Sculpin, and Cabezon are in gfsynopsis but have not been
designated an `iphc_common_name` in `iphc-spp-names.csv` (have to do that manually). Though Sandpaper Skate,
Salmon Shark, and Great Sculpin do show up has having IPHC data in 2019 gfsynopsis report,
but looks like only data from GFBio, looking carefully at the
`data_for_all_species` vignette for 2020:
http://htmlpreview.github.io/?https://github.com/pbs-assess/gfiphc/blob/master/vignettes/data_for_all_species.html
They did not have 2020 IPHC data, but do for 2021 (GS had 1995 and 1996 as zeros;
don't think others did). Cabezon has no previous data.

So, need to add those species to `iphc-spp-names.csv`, which may discover some
old data for those years when I redo the vignettes, as it seems strange that
they never seem to show up in the 20-hook-only data, just in GFBio.

Also add these to the `ignore_obvious` list in `check_iphc_spp_name()`:

"Sea Whip",
"Stylaster campylecus (coral)",
"Sun Sea Star",
"Jellyfish",
"Unident. Salmon",
"unident. organic matter",
"Dungeness Crab"

That list already had Sunflower Sea Star in it, presumably the same as Sun Sea
Star.

Then redoing those above commands with updated code gives this, where some
species are returned because they are not non-groundfish ones (or Brittle Star
or Glass Sponge which we also kept in the past) that we want to automatically ignore:
```{r updatedspnames}
check_iphc_spp_name(countData2021)
# That still retains some we may want to think about further at some point, but
#  these are all in the overall list for all years:
setdiff(check_iphc_spp_name(countData2021),
        check_iphc_spp_name())
check_iphc_spp_name()
```

## Save data sets

```{r savedata}
usethis::use_data(countData2021,
                  overwrite = TRUE)

usethis::use_data(setData2021,
                  overwrite = TRUE)
```

Add descriptions for new years in `R/data.R`.
