---
title: "Analyses for predators of Haida Gwaii Pacific Herring"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyses for predators of Haida Gwaii Pacific Herring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 8
)
```

This vignette shows how to extract and analyse the data for predators of Haida
Gwaii Pacific Herring, including lumping species together
(e.g. `skates_combined`). This information is for a DFO Ecosystem Approach to
Fisheries Management Case Study on ecoystem influences of Haida Gwaii Pacific
Herring.

The area of interest is the same as shown in the [Restricted
area](analysis_for_restricted_area.html) vignette, and the shapefile is
saved in the package as `HG_herring_pred_area`.

```{r setup, results = FALSE}
library(gfiphc)
path_data = "HG-herring-predators-data"        # folder to save extracted raw data
path_results = "HG-herring-predators-results"  # folder to save results
```

## Data for species of interest

These are the species (herring predators, excluding skate species) identified by
Jennifer Boldt and used at the end of `data_for_one_species` vignette (the final
code in that needs to be run first to cache the data).
```{r definesp}
sp_vec <- c("yelloweye rockfish",
            "arrowtooth flounder",
            "lingcod",
            "north pacific spiny dogfish",
            "pacific cod",
            "pacific halibut",
            "redbanded rockfish",
            "sablefish",
            "walleye pollock")
```

We want a general rockfish category and general skates one. I just picked
yelloweye and redbanded for rockfish for now.

## Skate species

Unidentified (species of) skates were enumerated in the early years, but then
declined dramatically and had zero counts in 2001, 2002 and 2013, and presumably
for 2003 onwards (but those data are in GFbio which would have to be checked --
seems clear that the identification protocol improved through time). So after
investigating this (in earlier versions of this vignette), am now including
unidentified skate in the species list.

Looking through 2019 synopsis report and the plots below (and a look at the
data), here are the decisions on including each skate species:

 - Aleutian Skate: Looks like only enumerated explicitly since 2007, 2018 map
   shows mean 0.31 fish/skate. Include.
 - Abyssal Skate - no catch, ignore.
 - Broad Skate - no catch, ignore.
 - Big Skate - looks like enumerated explicitly since 1998, and 2018 map shows
   mean 0.69 fish/skate. Include.
 - Roughtail Skate - looks like only caught in three or four years (mean +ve sets
   1/177). No catch in 2018. Include.
 - Sandpaper Skate - only shows up for a few years. Mean +ve sets 5/177, 2018
   mean 0.14 fish/skate. Include.
 - Longnose Skate - only shows up since 1998 (presumably unidentified
   beforehand), mean +ve sets 57/135, 2018 mean 0.96 fish/skate. Include.
 -  Alaska Skate - only showed up in five years, 2018 mean 0.14 fish/skate. Only
   plotted since 2003 (like Sandpaper). Include.

```{r skatedefn}
skate_sp <- c("aleutian skate",
              "big skate",
              "roughtail skate",
              "sandpaper skate",
              "longnose skate",
              "alaska skate",
              "unidentified skate")

skate_ignore <- c("abyssal skate",
                  "broad skate")
```

Cache the data once (this is not evaluated each time):
```{r get_skates, eval=FALSE}
cache_pbs_data_iphc(sp_vec,
                    path = path_data)
cache_pbs_data_iphc(skate_sp,
                    path = path_data)
cache_pbs_data_iphc(skate_ignore,
                    path = path_data)
```
## Individual skate species

Calculate series for each skate species in turn (saved as `species_name_area`)
and plot figures of Series E and F and longest from those. The fourth panel has
Series EF which is the longest time series possible; if it is blank then Series
E or F are the longest (and automatically identified). But this only happens for
the rare species, and so not when the skate species are combined.

```{r repeat, warning = FALSE, results = FALSE}
cached_results = TRUE   # whether or not to use already-saved results
for(sp in skate_sp){
  assign(sp_hyphenate(sp,
                      underscore = TRUE,
                      area = TRUE),
         iphc_get_calc_plot_area(sp,
                                 cached_data = TRUE,
                                 cached_results = cached_results,
                                 path_data = path_data,
                                 path_results = path_results))
}
```

## Unidentified skate

Above figure shows that for 1995 and 1996 there are 'Unident. skate', and then
for a few years after that (but at lower numbers). Can see from species figures
above that indeed no skate species were identified for those years. And it looks
like then big and longnose were identified but not others for a few years.

First looking into the raw data:

```{r unidentcheck}
# 1995
unique(countData1995$spNameIPHC)[grep("Skate", unique(countData1995$spNameIPHC))]
grep("skate", unique(countData1995$spNameIPHC)) # none
# And a manual check through confirms, that all we need is "unident. Skate"

#1996 to 2002 has "unident. Skate", but only up to 2000:
unique(data1996to2002$spNameIPHC)[grep("Skate",
                                       unique(data1996to2002$spNameIPHC))]
grep("skate", unique(data1996to2002$spNameIPHC)) # none
dplyr::filter(data1996to2002,
              spNameIPHC == "unident. Skate") %>%
  tail()

# And there were less unidentified through time:
dplyr::filter(data1996to2002,
              spNameIPHC == "unident. Skate") %>%
  dplyr::select(year) %>%
  table()


# 2013 (not in GFbio), no unident. Skate (checking 'skate' and 'Skate'):
unique(countData2013$spNameIPHC)[grep("Skate", unique(countData2013$spNameIPHC))]
grep("skate", unique(countData2013$spNameIPHC)) # none

# 2003-2019 (except 2013) - in GFbio
# See Issue #16 - not sure what the species code would be, but there may not be
# one if they were all identified after 2000.

# 2020 (not in GFbio), no unident. Skate (checking 'skate' and 'Skate'):
unique(countData2020$spNameIPHC)[grep("Skate", unique(countData2020$spNameIPHC))]
grep("skate", unique(countData2020$spNameIPHC)) # none
```

So, given none in 2001 or 2002 and numbers of unidentified declined from 1996 to
2000, it's probably the case that there were all identified from 2001
onwards. See Issue #16.

So, we can use `unidentified skate` as a species name (done above). Load in and
examine that data:
```{r unidentsummary}
unidentified_skate_area$sp_set_counts_with_area %>%
  dplyr::select(year) %>%
  table()
```
But there no positive counts after 2000, looking at first 20 hooks (Series E)
and then all hooks (Series F).
```{r unidentsummary2}
dplyr::select(unidentified_skate_area$ser_E_and_F$ser_E,
              year,
              num_pos20)
dplyr::select(unidentified_skate_area$ser_E_and_F$ser_F,
              year,
              num_pos)
```

## Merge together skate species

Now want to lump the skates in `skate_sp` together. Note that species counts get
named as `N_it.1` etc. While each one has NA's, none of the summed `N_it_sum` or
`N_it20_sum` do, because `get_combined_species()` ignores NA's when making the
sums. This is always recalculated (in case earlier cached data have changed).

```{r combine, warning = FALSE}
sp <- "skates combined"
skates_combined <- get_combined_species(sp_vec = skate_sp,
                                        path = path_data,
                                        save_RDS_name = sp_hyphenate(sp))
summary(skates_combined$set_counts)

# How many NA's in certain columns (none for N_it_sum or N_it20_sum
skates_combined$set_counts %>%
  dplyr::select(-c("station", "lat", "lon", "usable", "standard")) %>%
  dplyr::summarise_all( ~ sum(is.na(.x))) %>%
  t()
```

Note that by simply adding the counts of each skate species, we are
implicitly assuming that selectivity is the same across skate species. So true
changes in relative abundance may not be accurately reflected in changes in the
summed index.
<!-- Started an example: We are summing relative indices, so if one species (call it 'roller skate') is twice as
abundant as another (call it 'ice skate') but
has half the selectivity then their relative catch rates (fish per 100 hooks)
will be equal at, say, 1 fish per 100 hooks, given a summed rate of 2 fish per
100 hooks. But if roller skates double in abundance then ...  -->

Now use to calculate Series E, F and EF calculations for the restricted area:
```{r combineEF, warning = FALSE, results = FALSE}
assign(sp_hyphenate(sp, underscore = TRUE, area = TRUE),
       iphc_get_calc_plot_area(sp,
                               cached_data = TRUE,
                               cached_results = cached_results,
                               path_data = path_data,
                               path_results = path_results))
```


## Individual non-skate species

Similarly for the other species:
```{r showmult, warning = FALSE, results = FALSE}
for(sp in sp_vec){
  assign(sp_hyphenate(sp, underscore = TRUE, area = TRUE),
         iphc_get_calc_plot_area(sp,
                                 cached_data = TRUE,
                                 cached_results = cached_results,
                                 path_data = path_data,
                                 path_results = path_results))
}
```

## Format for Jennifer's EAFM summary

Save annual indices (with uncertainty) for all species of interest, in the
format for the HG herring predators' EAFM summary (first manually save previous
versions in `for-jen-earlier-versions/` if want to keep them).

```{r allsp}
all_sp_longest <- list()
sp_vec_and_combined_skates <- c(sp_vec, "skates combined")
# Do this to easily create arguments for iphc_format_for_EAFM(); can't easily
#  functionalise. Each element is names for the species, and is a tibble of
#  longest time series.
for(sp in sp_vec_and_combined_skates){
  all_sp_longest[[sp]] <- get(paste0(sp_hyphenate(sp,
                                                  underscore = TRUE,
                                                  area = TRUE)))$series_longest$ser_longest
}

eafm_res <- iphc_format_for_EAFM(sp_vec_and_combined_skates,
                                 path = path_results,
                                 all_sp_longest) # also saves .csv, and .rds of
                                                 #  all_sp_longest list

eafm_res
```

Note that the analyses may not yet fully account for all possibilities (for all
areas and species), particularly for rarer species. So do check that the output
is sensible by inspecting the above time series figures. And the analyses do not
(yet) include hook competition, which is currently being investigated.
