---
title: "Extract and analyse the IPHC data for a restricted area"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extract the IPHC data for one species}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

This vignette shows how to extract and analyse the data for a restricted area
(rather than the full coast). The example area is for predators of Haida Gwaii
Herring, and we'll stick with Yelloweye Rockfish as the example species (since
those data are saved in the package).

```{r setup, results = FALSE}
library(gfiphc)
library(dplyr)   # TODO hopefully not need?
```

## Data for species of interest

Define common species name:
```{r species}
sp <- "yelloweye rockfish"
sp_short_name <- "Yelloweye"  # short name for legends
```

As for the `data_for_one_species.Rmd` vignette, change the above names to your
species of interest and have someone at PBS run the next chunk of code to extract data from
the GFBio database.
```{r extract, eval = FALSE}
# This chunk will only work within PBS
cache_pbs_data_iphc(sp)
```

Once an external person has the `species-name.rds` file, the two lines in the
next chunk can be appropriately commented/uncommented to load the data. For the
vignette to always run we will stick with the Yelloweye Rockfish data.
```{r loadrds}
# Comment this out to analyse non-yelloweye data:
sp_set_counts <- yelloweye_rockfish

# Uncomment this to analyse non-yelloweye data:
# sp_set_counts <- readRDS(paste0(gsub(" ", "-", sp), ".rds"))
```

## Analyses for subset of stations

The area of interest is saved in `gfiphc` as `HG_herring_pred_area`:
```{r predators, results = FALSE}
plot_BC(main = "All 2008 stations")

PBSmapping::addPolys(HG_herring_pred_area,
                     col = "pink")

plot_iphc_map(sp_set_counts$set_counts,
              sp = NULL,
              years = 2008,
              lat_cut_off = NULL,
              add_to_existing = TRUE)
```

Create tibble of set counts with extra column `in_area`, which is `TRUE/FALSE`
whether or not that set is within the area of interest:
```{r inarea, results = FALSE}
sp_set_counts_with_area <- add_in_area(sp_set_counts$set_counts,
                                       area = HG_herring_pred_area)
```

Each catch count now also has an `in_area` logical attribute:
```{r inareares}
dplyr::select(sp_set_counts_with_area, year, station, lat, lon, E_it, C_it, usable,
              standard, in_area)
```

Colour code on map:
```{r area, results = FALSE}
plot_BC(main = "All 2008 stations")

PBSmapping::addPolys(HG_herring_pred_area,
                     col = "pink")

plot_iphc_map(sp_set_counts_with_area,
              sp = NULL,
              years = 2008,
              lat_cut_off = NULL,
              add_to_existing = TRUE,
              indicate_in_area = TRUE)
```

2018 had expanded (non-standard) stations, which are shown as open circles:
```{r area2018, results = FALSE}
plot_BC(main = "All 2018 stations")

PBSmapping::addPolys(HG_herring_pred_area,
                     col = "pink")

plot_iphc_map(sp_set_counts_with_area,
              sp = NULL,
              years = 2018,
              lat_cut_off = NULL,
              add_to_existing = TRUE,
              indicate_in_area = TRUE)
```

Make a movie to check each year:
```{r, animation.hook = 'gifski', interval = 1.5, fig.height = 6, results = FALSE}
for(i in unique(sp_set_counts_with_area$year)){
  plot_BC(main = paste0("All ", i, " stations"))

  PBSmapping::addPolys(HG_herring_pred_area,
                       col = "pink")

  plot_iphc_map(sp_set_counts_with_area,
                sp = NULL,
                years = i,
                lat_cut_off = NULL,
                add_to_existing = TRUE,
                indicate_in_area = TRUE)
}
```

## Calculate longest time series possible for the area of interest

In Appendix G of the synopsis report we described analyses to calculate as long
a time series as possible by combining data from years when only the first 20 hooks
were enumerated for each skate (Series A) and from years when all hooks could be enumerated
from each skate (Series B); there are overlapping years when all hooks were
enumerated and the data are available at the hook-by-hook level. The resulting
Series AB applied to only north of WCVI, but was tested to see if it was
representative of the whole coast.

For stations restricted to the area of interest, here we define Series E to be
based on the first 20 hooks from each skate, and Series F to be based on all
hooks from each skate. We do a similar analysis as described in the synopsis
report to generate the longer Series EF (or just one of Series E and F if there are
insufficient data). It is up to the user to generate maps (or movie) of their area
of interest to confirm that it makes sense to use all stations with the
area. For example, considering an area that reaches further down the west coast
of Vancouver Island would not be appropriate since it was not surveyed in some
years (Table G.2 of synopsis report).

The calculations are as follows, yielding figure with four plots similar to Figures G.1 and
G.2 of the synopsis report (that were for Series A and B):
```{r serEF}
ser_E_and_F <- calc_iphc_ser_E_and_F(sp_set_counts_with_area)
ser_E_and_F
```

Then to calculate the longest possible time series:
```{r serlongest, fig.width = 10, fig.height = 8}
series_longest <- calc_iphc_ser_EF(ser_E_and_F)
series_longest

plot_IPHC_ser_four_panels(ser_E_and_F, series_longest)
```

Note that the code may not yet fully account for all possibilities (for all
areas and species), particularly for rarer species. So do check that the output
is sensible.
