---
title: "Extract the IPHC data for one species"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extract the IPHC data for one species}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

```{r setup}
library(gfiphc)
```

<!-- Adapting code written by Andrew Edwards for Beau Doherty at Landmark -->
<!-- Fisheries in March 2019 and updated in September 2019.-->
<!-- Also using MEE_reproduce_2.Rmd vignette from sizeSpectra as a template -->

## Data for all skates and stations, and counts for one species

Extract and save the IPHC data for a single species, using Yelloweye Rockfish as an
example, defined here:

```{r species}
sp <- "yelloweye rockfish"
```

Someone at PBS has to run the first chunk of code to extract data from
the GFBio database and send the external collaborator the resulting four .rds
files. The remaining code will run outside of PBS. Within PBS, the first chunk doesn't need
to be re-run once the .rds files have been saved locally.

```{r extract}
# This chunk will only work within PBS
cache_pbs_data_iphc(sp)
                                       # That creates yelloweye-rockfish.rds. [This is
                                       #  what gfsynopsis::get_data_iphc() calls,
                                       #  via gfsynopsis::get_data() in
                                       #  report/make.R of gfsynopsis repo.]

cache_pbs_data_iphc("hook with bait")
                                       # creates hook-with-bait.rds, same as
                                       # sp_set_counts but for 'species' hook
                                       # with bait; needed if looking into hook competition.

saveRDS(get_iphc_sets_info(), "sets-other-years.rds", compress = TRUE)
                                       # extracts the set level data from
                                       #  GFBio for 2003 onwards (excluding 2013)

saveRDS(get_iphc_skates_info(), "skates-other-years.rds", compress = TRUE)
                                       # extracts skate level data from GFBio
                                       #  for 2003 onwards (excluding 2013)
```

Once an external person has the four .rds files, the rest of this code will run
outside of PBS.
```{r loadrds}
sp_set_counts <- readRDS(paste0(gsub(" ", "-", sp), ".rds"))
                                       # For each set, contains the various
                                       # catch rates for this species, the lat
                                       # and lon, and whether the set
                                       # is usable or not. Includes the basic
                                       # information for each set, but further
                                       # details are available in the
                                       # objects described below
hooks_with_bait <- readRDS("hook-with-bait.rds")
sets_other_years <- readRDS("sets-other-years.rds")
skates_other_years <- readRDS("skates-other-years.rds")
```


Other data sets are already built into the gfiphc package, the main ones being:
```{r builtin}
setData1995      # Locations and effective skate values of stations for the 1995 survey
countData1995    # Counts of each species at each station in 1995
data1996to2002   # Catches at each station from 1996 to 2002
setData2013      # Station details for 2013 survey
setDataExpansion # Station details regarding expansion stations from 2018 (and
                 # possibly onwards)
```
See `?<dataset>` for details of each, and `data(package = "gfiphc")` for other
data sets. Notation such as `E_it20` matches the write-up in the Groundfish
Synopsis report.

The formats are different to each other due to the way the data are
available. For example, there are no `hooksObserved` values available for 1995
and 2013, which will complicate consideration of hook competition
for those years. `setDataExpansion` is needed to identify stations in the expanded
grid (that were not fished in previous years).

Look at the data for the species of interest:
```{r load}
sp_set_counts
                  # For each set, the calculatable catch rates for that
                  # species, plus lat and lon and whether the set is usable.
                  # See ?get_all_iphc_set_counts for full details.

summary(sp_set_counts$set_counts)

sets_each_year <- sp_set_counts$set_counts %>% group_by(year) %>%
  summarise(total = n())
plot(sets_each_year$year, sets_each_year$total, type = "o",
     xlab = "Year", ylab = "Sets each year")
```

This shows that the number of sets changes over time. In particular, there were
lots more stations in 2018 and these need to be accounted for when calculating
the catch rates. THEY HAVE NOT YET BEEN IN THE AUTOMATED FUNCTIONS ABOVE. Also
were not accounted for in the gfsynopsis report (when we added in the 2018 data
quite late, see Issue #14, which notes that the latest Yelloweye Rockfish
assessment properly accounted for this).

Need to look at locations of stations, and check in detail what happens for 2018
onwards.

## Locations of stations

Will adapt old code to create maps of stations, but I think Beau at Landmark
does this anyway, and correctly ignored the new stations in 2018.

## Not running yet, can ignore for now:

Robyn did this for P Cod (but will not get final year(s) right since I haven't yet
corrected for changing stations) to take a quick look at the data, but didn't
end up using it.

```{r pcod, eval = FALSE}
iphc_raw <- gfiphc::tidy_iphc_survey(
  get_iphc_hooks("pacific cod"),
  get_iphc_skates_info(),
  get_iphc_sets_info()
)
iphc_index <- gfiphc::calc_iphc_full_res(iphc_raw)
```


The code used in `gfsynopsis/R/make-pages.R` is the following, simplifying it
here to then check that it matches what is done above (not finished checking yet):
```{r gfsynopsis, eval = FALSE}
#if (!is.null(dat_iphc)) {
#    if (!file.exists(iphc_index_cache_spp)) {
      iphc_set_counts_sp <-
        gfiphc::calc_iphc_full_res(dat_iphc$set_counts)

      saveRDS(iphc_set_counts_sp, file = iphc_index_cache_spp, compress = FALSE)
    } else {
      iphc_set_counts_sp <- readRDS(iphc_index_cache_spp)
    }

    iphc_set_counts_sp_format <- tryCatch({
      gfiphc::format_iphc_longest(iphc_set_counts_sp)
    }, error = function(e) NA)
    # Remove existing (GFbio) based IPHC series with longer ones from new calcs
    if (!is.na(iphc_set_counts_sp_format)) {
      dat_tidy_survey_index <- dat_tidy_survey_index %>%
        filter(survey_abbrev != "IPHC FISS") %>%
        rbind(iphc_set_counts_sp_format)
    }
  }
```
