---
title: "Extract the IPHC data for one species"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extract the IPHC data for one species}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

```{r setup}
library(gfiphc)
```

<!-- Adapting code written by Andrew Edwards for Beau Doherty at Landmark -->
<!-- Fisheries in March 2019 and updated in September 2019.-->
<!-- Also using MEE_reproduce_2.Rmd sizeSpectra vignette as a template -->

## Data for all skates and stations, and counts for one species

Extract and save the IPHC data for a single species, using Yelloweye Rockfish as an
example. Someone at PBS has to run the first chunk of code to extract data from
the GFBio database and send the external collaborator the resulting four .rds
files. The rest of the code will run outside of PBS.

```{r extract}
# Will only run at PBS as accesses the GFBio database
sp <- "yelloweye rockfish"
cache_pbs_data_iphc(sp)                # creates yelloweye-rockfish.rds. This is
                                       # what gfsynopsis::get_data_iphc() calls,
                                       # via gfsynopsis::get_data() in
                                       # report/make.R of gfsynopsis repo.

cache_pbs_data_iphc("hook with bait")  # creates hook-with-bait.rds

sets <- get_iphc_sets_info()           # extracts the set level data from
                                       #  GFBio for 2003 onwards (excluding 2013)
saveRDS(sets, "sets.RDS", compress = TRUE)

skates <- get_iphc_skates_info()       # extracts skate level data from GFBio
                                       #  for 2003 onwards (excluding 2013)
saveRDS(skates, "skates.RDS", compress = TRUE)
```

Other data sets are already built into the gfiphc package, the main ones being:
```{r builtin}
setData1995      # Locations and effective skate values of stations for the 1995 survey
countData1995    # Counts of each species at each station in 1995
data1996to2002   # Catches at each station from 1996 to 2002
setData2013      # Station details for 2013 survey
```
See `?<dataset>` for details of each, and `data(package = "gfiphc")` for other
data sets. Notation such as `E_it20` matches the write-up in the Groundfish
Synopsis report.

The formats are different to each other due to the way the data are
available. For example, there are no `hooksObserved` values available for 1995
and 2013, which will complicate consideration of hook competition
for those years.

Need to look at locations of stations, since in 2019 they expanded again.

## Locations of stations









Robyn did for P Cod (but will not get final year(s) right since I haven't yet
corrected for changing stations).

```{r pcod}
iphc_raw <- gfiphc::tidy_iphc_survey(
  get_iphc_hooks("pacific cod"),
  get_iphc_skates_info(),
  get_iphc_sets_info()
)
iphc_index <- gfiphc::calc_iphc_full_res(iphc_raw)
```


The code used in `gfsynopsis/R/make-pages.R` is the following, simplifying it
here to then check that it matches what is done above (not finished checking yet):
```{r gfsynopsis, eval = FALSE}
#if (!is.null(dat_iphc)) {
#    if (!file.exists(iphc_index_cache_spp)) {
      iphc_set_counts_sp <-
        gfiphc::calc_iphc_full_res(dat_iphc$set_counts)

      saveRDS(iphc_set_counts_sp, file = iphc_index_cache_spp, compress = FALSE)
    } else {
      iphc_set_counts_sp <- readRDS(iphc_index_cache_spp)
    }

    iphc_set_counts_sp_format <- tryCatch({
      gfiphc::format_iphc_longest(iphc_set_counts_sp)
    }, error = function(e) NA)
    # Remove existing (GFbio) based IPHC series with longer ones from new calcs
    if (!is.na(iphc_set_counts_sp_format)) {
      dat_tidy_survey_index <- dat_tidy_survey_index %>%
        filter(survey_abbrev != "IPHC FISS") %>%
        rbind(iphc_set_counts_sp_format)
    }
  }
```
